Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-07 papers/static-checker/ecoop2024/ecoop2024.tex
update dsl section

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, failwith (\StackOne) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\section {Domain Specific Language}
Michelson, being a stack-based language, may be less familiar to many developers compared to popular languages like Python, Java, or C. To bridge this gap, several high-level languages such as LIGO, SmartPy, and Liquidity have been developed. These languages enable developers to write Tezos smart contracts in a more readable manner, resembling popular programming languages. Contracts written in these languages are then compiled down to Michelson.

In addition, Tezos users, who don't have much knowleage about Michelson language may wish to verify smart contract code before deployment. Our framework aims to provide a tool that doesn't require users to be Michelson experts. We intend to design a domain-specific language that enables users to specify smart contract properties in a mathematical way, making it readable and easy  to understand.

Given that our framework utilizes an SMT solver, specifically Z3, for property verification, we offer a language that allows users to express formulas compatible with the capabilities of the SMT solver (Z3). This empowers users to write formulas that the solver can effectively handle for property verification.

One of the main challenges with formal verification is that it ofen requires the specifier to express the program code in a specification language. This poses two problems: it demands expertise in formal verification from the specifier and introduces a gap between the actual code and its specification. In our framework, this gap is eliminated as the exact Michelson code used for verification precisely corresponds to the real code. Users are not required to be formal verification experts. The necessary Michelson knowledge revolves around understanding data types and their usage when calling a smart contract on the Tezos blockchain, both as parameters and in storage.

>>>>

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, failwith (\StackOne) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\section {Domain Specific Language}
Michelson, as a stack-based language, may pose a challenge for developers less familiar with it compared to more mainstream languages like Python, Java, or C. To address this, several high-level languages such as LIGO, SmartPy, and Liquidity have been created. These languages provide a more readable syntax, resembling popular programming languages, allowing developers to write Tezos smart contracts more comfortably. These high-level contracts are then compiled down to Michelson.

For Tezos users who may lack expertise in Michelson, there is a desire to verify smart contract code before interaction. Our framework addresses this by introducing a domain-specific language (DSL) that facilitates the specification of smart contract properties in a mathematical and readable format. The goal is to create a language that is accessible and easily understandable, even for users who are not Michelson experts.

To achieve this, our framework employs an SMT solver, specifically Z3, for property verification. The DSL is designed to allow users to express formulas that align with the capabilities of the SMT solver (Z3). This empowers users to write formulas that the solver can effectively handle, facilitating property verification in a more user-friendly manner.

One of the main challenges with formal verification is that it ofen requires the users to specify the system in a specification language. This poses two problems: it demands expertise in formal verification from the specifier and introduces a gap between the actual code and its specification. In our framework, this gap is eliminated as the exact Michelson code used for verification precisely corresponds to the real code. Users are not required to be formal verification experts. The necessary Michelson knowledge revolves around understanding data types and their usage when calling a smart contract on the Tezos blockchain, both as parameters and in storage.
